Malacology: A Programmable Storage System

### 2016-10-11 paper/paper.tex
paper: more jeff edits for motivation/intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{abstract}
Storage systems are caught between rapidly changing data
processing systems and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to adapt both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows unprecedented
flexibility for storage systems to evolve without sacrificing the robustness of
its code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by constructing two services out of existing abstractions: a
file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to make trade-offs for lease capabilities, balance load, propogate
interfaces, and recover from failure.
\end{abstract}

@@

A storage system implements abstractions designed to persistently store data
and must exhibit a high level of correctness to prevent data loss.  Storage
systems have evolved around storage devices that often were orders of magnitude
slower than CPU and memory, and therefore could dominate overall performance if
not used carefully. Over the last few decades members of the storage systems
community have developed ingenious strategies to meet correctness requirements
while somewhat hiding the latency of traditional storage
media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
of storage systems have preferred systems with highly standardized APIs and
lowest common denominator abstract data types such as blocks of bytes and byte
stream files~\cite{armbrust_view_2010}.

@@

In this paper we investigate an approach that focuses on generalizing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By doing so one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency. We will refer to this programmability as
\emph{programmable storage}, which differs from \emph{active storage} (the
injection and execution of arbitrary code in a storage system) and
\emph{software-defined storage} (the control of thin-provisioning of storage).

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system for constructing new
services by re-purposing existing subsystem abstractions of the storage stack.
We build the framework in Ceph, a popular, open source software storage stack,
by leveraging a broad spectrum of existing services, including distributed
locking and caching services provided by metadata servers, durability and
logical storage devices provided by the backend object store, and propagation
of consistent cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, this framework is
expressive enough to provide the functionality necessary for implementing new
services. Our contributions are:

\paragraph*{A programmable storage system} that re-uses and extends existing
abstractions and services of the underlying storage stack. It
includes the following building blocks:

@@

\paragraph*{Two distributed services} that demonstrate the feasibility of using
the programmable storage approach:

@@

\item A re-implementation of the programmable Mantle metadata load balancing
service~\cite{sevilla:sc15-mantle} using Malacology

>>>>

\begin{abstract}
Storage systems are caught between rapidly changing data
processing systems and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to adapt both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows a large degree of
flexibility for storage systems to evolve without sacrificing the robustness of
its code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by composing existing primitives into two new higher-level
services: a file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to make trade-offs for latency/throughput using lease capabilities, balance load, propogate
interfaces, and recover from failure.
\end{abstract}

@@

A storage system implements abstractions designed to persistently store data
and must exhibit a high level of correctness to prevent data loss.  Storage
systems have evolved around storage devices that often were orders of magnitude
slower than CPU and memory, and therefore could dominate overall performance if
not used carefully. Over the last few decades members of the storage systems
community have developed clever strategies to meet correctness requirements
while somewhat hiding the latency of traditional storage
media~\cite{brewer_disks_2016}. To avoid lock-in by a particular vendor, users
of storage systems have preferred systems with highly standardized APIs and
lowest common denominator abstract data types such as blocks of bytes and byte
stream files~\cite{armbrust_view_2010}.

@@

In this paper we investigate an approach that focuses on generalizing and
composing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstractions, one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency. We will refer to this programmability as
\emph{programmable storage}, which differs from \emph{active storage} (the
injection and execution of arbitrary code in a storage system) and
\emph{software-defined storage} (the control of thin-provisioning of storage).

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system for constructing new
services by re-purposing existing subsystem abstractions of the storage stack.
We build the framework in Ceph, a popular, open source software storage stack.
We choose Ceph to demonstrate the concept of programmable stoarge because it
offers a broad spectrum of existing services, including distributed locking and
caching services provided by metadata servers, durability and logical storage
devices provided by the backend object store, and propagation of consistent
cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, this framework is
expressive enough to provide the functionality necessary for implementing new
services. Our contributions are:

{\it \textbf{We define a programmable storage system}} to be a storage system that re-uses and extends existing
abstractions and services of the underlying software stack. Our prototype
includes (but is not limited to) the following building blocks:

@@

{\it \textbf{We implement two distributed services}} using programmable storage
to demonstrate its feasability:

@@

\item An implementation of the programmable Mantle metadata load balancing
service~\cite{sevilla:sc15-mantle}
